Revealing transient strain in geodetic data with Gaussian process regression

### 2017-10-30 RevisedSubmission/hines_hetland_2017_gji.tex
did more

We next establish our prior model for transient displacements.
Specifically, we discuss our choice for the functions $X_i$ and $T_i$.
For $X_i$, we use the squared exponential (SE) covariance function,
\begin{equation}\label{eq:SE}
X_i(\pos,\pos') = \exp\left(\frac{-||\pos - \pos'||_2^2}{2 \ell^2}\right).
\end{equation}
The SE covariance function is commonly used for kriging
\citep[e.g,][]{Cressie1992} and Gaussian process regression
\citep[e.g.,][]{Rasmussen2006}. In terms of geodetic applications,
\citet{Kato1998} and \cite{El-Fiky1999} demonstrated that the SE
covariance function accurately describes the spatial covariance of
secular GNSS derived velocities in Japan.

@@

We use the REML method to determine the appropriate values for the
parameters in $X_i$ and $T_i$, $\mitbf{\theta} = \{\ell, \phi, \tau
\}$. We divide the GNSS data into seven subsets that are four months
long and each centered on the time of a SSE. The times of the seven
SSEs are determined with tremor records from \cite{Wech2010}. We
choose to make each data subsets four months long because it is long
enough to encompass a SSE in the Pacific Northwest, while it is still
computationally tractable to estimate $\mitbf{\theta}$. However, four
months is too short to resolve the sinusoids in $\data_i$, and they
are omitted from $\data_i$ in this REML analysis. We find the optimal
$\mitbf{\theta}$ for each subset of data, for each displacement
component, and for each choice of $T_i$. The REML likelihood function
that we are maximizing with respect to $\mitbf{\theta}$ is now
\begin{equation}\label{eq:REMLPrior} 
\mathcal{L}(\mitbf{\theta}) =
\left(\frac{\left| \G^T \G \right|}{(2\pi)^{n-6m}
                                    \left| \mitbf{\Sigma}(\mitbf{\theta}) \right| 
                                    \left| \G^T \mitbf{\Sigma}(\mitbf{\theta})^{-1} \G \right|}\right)^{\frac{1}{2}} 
                              e^{-\tfrac{1}{2} \data_i^{*T} \mitbf{K}(\mitbf{\theta}) \data_i^*},
\end{equation}
where
\begin{equation}
\mitbf{K}(\mitbf{\theta}) = \mitbf{\Sigma}(\mitbf{\theta})^{-1} - 
                            \mitbf{\Sigma}(\mitbf{\theta})^{-1} \G
                            \left( \G^T \mitbf{\Sigma}(\mitbf{\theta})^{-1} \G \right)^{-1}
                            \G^T \mitbf{\Sigma}(\mitbf{\theta})^{-1}
\end{equation}
and $\mitbf{\Sigma}(\mitbf{\theta}) = \mitbf{C}_{\eta_i} +
C_{u_i}(\points,\points;\mitbf{\theta})$. In the above equation,
$\data^*_i$ consists of a single subset of data. The estimated
parameters are summarized in Table 1. Based on the interquartile
ranges, the estimated parameters for the SE and Wendland covariance
functions do not vary significantly between SSEs. This suggests that
the median of the estimated parameters should be appropriate for all
the Pacific Northwest SSEs. For the IBM model, there are several
anomalously large values for $\ell$ and $\phi$, which is why the
interquartile ranges are large.

>>>>

We next establish our prior model for transient displacements.
Specifically, we discuss our choice for the spatial and temporal
functions making up $C_{u_i}$, $X_i$ and $T_i$. For $X_i$, we use the
squared exponential (SE) covariance function,
\begin{equation}\label{eq:SE}
X_i(\pos,\pos') = \exp\left(\frac{-||\pos - \pos'||_2^2}{2 \ell^2}\right).
\end{equation}
The SE covariance function is commonly used for kriging
\citep[e.g,][]{Cressie1992} and Gaussian process regression
\citep[e.g.,][]{Rasmussen2006}. In terms of geodetic applications,
\citet{Kato1998} and \cite{El-Fiky1999} demonstrated that the SE
covariance function accurately describes the spatial covariance of
secular GNSS derived velocities in Japan.

@@

We use the REML method to determine the appropriate values for the
parameters in $C_{u_i}$, $\mitbf{\theta} = \{\ell, \phi, \tau \}$.
Just as in Section \ref{sec:NoiseModel}, we pick $\mitbf{\theta}$ to
maximize the probability of drawing $\data_i^*$ from $\data_i$, but
now $\data_i$ includes $u_i(\points)$, and $\mitbf{\theta}$ refers to
the parameters for $C_{u_i}(\points,\points)$ rather than
$\mitbf{C}_{\eta_i}$. We divide the GNSS data into seven subsets that
are four months long and each centered on the time of a SSE. The times
of the seven SSEs are determined with tremor records from
\cite{Wech2010}. We find the optimal $\mitbf{\theta}$ for each subset
of data, for each displacement component, and for each choice of
$T_i$. We choose to make each data subsets four months long because it
is long enough to encompass a SSE in the Pacific Northwest, while it
is still computationally tractable to estimate $\mitbf{\theta}$. The
REML likelihood function that we are maximizing with respect to
$\mitbf{\theta}$ is now
\begin{equation}\label{eq:REMLPrior} 
\mathcal{L}(\mitbf{\theta}) =
\left(\frac{\left| \G^T \G \right|}{(2\pi)^{n-6m}
                                    \left| \mitbf{\Sigma}(\mitbf{\theta}) \right| 
                                    \left| \G^T \mitbf{\Sigma}(\mitbf{\theta})^{-1} \G \right|}\right)^{\frac{1}{2}} 
                              e^{-\tfrac{1}{2} \data_i^{*T} \mitbf{K}(\mitbf{\theta}) \data_i^*},
\end{equation}
where
\begin{equation}
\mitbf{K}(\mitbf{\theta}) = \mitbf{\Sigma}(\mitbf{\theta})^{-1} - 
                            \mitbf{\Sigma}(\mitbf{\theta})^{-1} \G
                            \left( \G^T \mitbf{\Sigma}(\mitbf{\theta})^{-1} \G \right)^{-1}
                            \G^T \mitbf{\Sigma}(\mitbf{\theta})^{-1}
\end{equation}
and $\mitbf{\Sigma}(\mitbf{\theta}) = \mitbf{C}_{\eta_i} +
C_{u_i}(\points,\points;\mitbf{\theta})$. In the above equation,
$\data^*_i$ consists of a single subset of data. The estimated
parameters are summarized in Table 1. Based on the interquartile
ranges, the estimated parameters for the SE and Wendland covariance
functions do not vary significantly between SSEs. This suggests that
the median of the estimated parameters should be appropriate for all
the Pacific Northwest SSEs. For the IBM model, there are several
anomalously large values for $\ell$ and $\phi$, which is why the
interquartile ranges are large.
